---
title: "Weekend Crypto Price Changes"
author: "Andrew Alexandru"
date: "10/6/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
gemini_BTCUSD_day <- read_csv("C:/Users/Andrew/OneDrive/Desktop/gemini_BTCUSD_day.csv", 
                              col_types = cols(Date = col_character()), 
                              skip = 1)
gemini_ETHUSD_day <- read_csv("C:/Users/Andrew/OneDrive/Desktop/gemini_ETHUSD_day.csv", 
                              col_types = cols(Date = col_character()), 
                              skip = 1)




#BTC cleaning
btcdata = data.frame(gemini_BTCUSD_day$Date, gemini_BTCUSD_day$Open, gemini_BTCUSD_day$Close)
colnames(btcdata) = c("Date","Open","Close")
btcdata$Change = ((btcdata$Close-btcdata$Open)/btcdata$Open)*100
btcdatesplit = data.frame(do.call('rbind', strsplit(as.character(btcdata$Date),' ',fixed=TRUE)))
btcdata$Day = weekdays(as.Date(btcdatesplit$X1,"%Y-%m-%d"))
btcdata = head(btcdata, -1)

#ETH cleaning
ethdata = data.frame(gemini_ETHUSD_day$Date, gemini_ETHUSD_day$Open, gemini_ETHUSD_day$Close)
colnames(ethdata) = c("Date","Open","Close")
ethdata$Change = ((ethdata$Close-ethdata$Open)/ethdata$Open)*100
ethdatesplit = data.frame(do.call('rbind', strsplit(as.character(ethdata$Date),' ',fixed=TRUE)))
ethdata$Day = weekdays(as.Date(ethdatesplit$X1,"%Y-%m-%d"))
ethdata = head(ethdata, -1)
```

## Do Crypto Prices Increase on the Weekends

With the rise of easy to use trading apps, are retail investors increasing the prices of coins during the weekend, when they have more time?


## BTC Daily Price Changes
```{r echo=TRUE}
hist(btcdata$Change)
```

## Normality Test for BTC Price Changes

```{r echo=TRUE}
shapiro.test(btcdata$Change)
```

## QQ Plot for BTC Price Changes
```{r echo=TRUE}
qqnorm(btcdata$Change)
```

## Does BTC's price increase on weekends?
```{r}
btchighersuc = nrow(btcdata[btcdata$Day %in% c("Sunday","Saturday") & btcdata$Change > 0 ,])
btchighertotal = nrow(btcdata[btcdata$Day %in% c("Sunday","Saturday"),])
binom.test(btchighersuc,btchighertotal, alternative = "greater")
```
Alt: The probability that BTC daily return will be positive on a weekend day is greater then 50%.

Null: The probability that BTC daily return will be positive on a weekend day is less then or equal to 50%

## Yes!
Because it increases in general.

## BTC price increases everyday
```{r}
btchighersuctotal = nrow(btcdata[btcdata$Change>0,])
binom.test(btchighersuctotal,nrow(btcdata), alternative = "greater")
```
Alt: The probability that BTC daily return will be positive on an given day is greater then 50%.

Null: The probability that BTC daily return will be positive on an given day is less then or equal to 50%

## The same is true for ETH
```{r}
ethhighersuc = nrow(ethdata[ethdata$Day %in% c("Sunday","Saturday") & ethdata$Change > 0 ,])
ethhighertotal = nrow(ethdata[ethdata$Day %in% c("Sunday","Saturday"),])
binom.test(ethhighersuc,ethhighertotal, alternative = "greater")
```
Alt: The probability that ETH daily return will be positive on a weekend day is greater then 50%.

Null: The probability that ETH daily return will be positive on a weekend day is less then or equal to 50%

---

```{r}
ethhighersuctotal = nrow(ethdata[ethdata$Change>0,])
binom.test(ethhighersuctotal,nrow(ethdata), alternative = "greater")
```
Alt: The probability that ETH daily return will be positive on any day is greater then 50%.

Null: The probability that ETH daily return will be positive on any day is less then or equal to 50%

## Yes, Crypto Prices Increase on the Weekends*
Because crypto prices increase in general.

Perhaps all the extra time trading can also be selling therfor increase volatility?

## BTC Volatility on Weekends
```{r}
btcvolsuc = nrow(btcdata[btcdata$Day %in% c("Sunday","Saturday") & abs(btcdata$Change) > 5 ,])
btcvoltotal = nrow(btcdata[abs(btcdata$Change) > 5 ,])
binom.test(btcvolsuc,btcvoltotal,p=2/7,alternative = "two.sided")
```
Alt: The probability of a greater then 5% swing occurring on the weekend is not 2/7.

Null: The probability of a greater then 5% swing occurring on the weekend is 2/7.


## ETH Volatility on Weekends
```{r}
ethvolsuc = nrow(ethdata[ethdata$Day %in% c("Sunday","Saturday") & abs(ethdata$Change) > 5 ,])
ethvoltotal = nrow(ethdata[abs(ethdata$Change) > 5 ,])


binom.test(ethvolsuc,ethvoltotal,p=2/7,alternative = "two.sided")
```
Alt: The probability of a greater then 5% swing occurring on the weekend is not 2/7.

Null: The probability of a greater then 5% swing occurring on the weekend is is 2/7.


## Difference of results
Possible explanations for my analysis:

- My definition of a weekend is a little odd as the data I used it at daily intervals at 4am.
- My definition of volatility is quite rudimentary as 5% swings or greater.

## Credits
My data is from: https://www.cryptodatadownload.com/data/gemini/

Seth, S. (2021, August 24). Bitcoin: Biggest price swings happen on weekends. Investopedia. Retrieved October 6, 2021, from https://www.investopedia.com/news/bitcoin-biggest-price-swings-happen-weekends/. 
