---
title: "Stat 523 Project 3"
author: "Andrew Alexandru"
date: "30065419"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

depth = c(9.3,9.4,9.6,10.0,9.4,9.3,9.8,9.9,9.2,9.4,9.5,9.7,9.7,9.6,10.0,10.2)
tips = rep(c("Tip1","Tip2","Tip3","Tip4"), each=4)
metal = rep(1:4, 4)

data = data.frame(depth,tips,metal)
datacharts = reshape(data = data, idvar="metal", v.names = "depth", timevar = "tips", direction = "wide")
datapresent = datacharts
colnames(datapresent) = c("Metal Ingot", "Tip 1", "Tip 2", "Tip 3", "Tip 4")
```

## Which Metal Cutting Tip is the Hardest?
> - Testing out 4 different kinds of tips
> - Testing on 4 different metal ingots (same metal but each ingot could be created at different temperatures resulting in different properties)
> - Depth will be in mm the length in which a given tip can penetrate the metal ingot at a specific pressure.

## RCBD Data
```{r}
datapresent
```
Data can be found at: https://web.ma.utexas.edu/users/mks/384E09/rcbdexampleslides.pdf

## Boxplots showing depth (mm) by Tip
```{r}
boxplot(datacharts$depth.Tip1,datacharts$depth.Tip2,datacharts$depth.Tip3,datacharts$depth.Tip4)
```

## Quade Test
Less then 5 treatment so we will do a Quade Test over the Friedman Test
```{r, echo=TRUE}
quade.test(depth, groups = tips, blocks = metal, p.adjust.methods = "none")
```

## Posthoc Quade Test
```{r, echo=TRUE, warning=FALSE}
library(PMCMRplus)
quadeAllPairsTest(y=depth, groups = tips, blocks = metal, p.adjust.method = "none")
```

## Comparison to parametric method?
```{r}
shapiro.test(datacharts$depth.Tip1)
shapiro.test(datacharts$depth.Tip2)
shapiro.test(datacharts$depth.Tip3)
shapiro.test(datacharts$depth.Tip4)
```

## Comparison to Pairwise Wilcox
```{r, echo=TRUE}
pairwise.wilcox.test(depth,tips, p.adjust.method = "none", exact=FALSE)
```

## Comparison
The pairwise posthoc Quade test gave us a p-value of **0.0047** between Tip 3 and Tip 4, while the pairwise Wilcox test only gave us **0.081**.

The variance we were able to explain away due to blocking the metal variable allows us to find a statistically significant result.

## Conclusion and Questions
Tip 4 is significantly harder then Tip 3

